SMT-Boosted Security Types for Low-Level MPC

### 2025-10-30 toplas26/lang.tex
checkpoint

>>>>

\paragraph{Sanity conditions on protocols} To simplify
the presentation we will disregard protocols that are semantically
ill-defined. That is, we generally assume that for any
configuration $(\store, \prog)$ there exists $\store$
with $(\store, \prog) \redxs \store$. This rules out
use of undefined variables and, in the adversarial semantics,
programs consisting of just a failing $\assert$.

### 2025-10-30 toplas26/smt.tex
checkpoint

\section{$\minicat$ Constraint Verification}
\label{section-smt}

In previous work it was shown that the semantics of $\minicat$ in
$\mathbb{F}_2$ can be implemented with Datalog
\cite{skalka-near-ppdp24}, which is a form of constraint
programming. In this paper we extend this idea to arbitrary prime
fields by using a more general form of SMT constraint programming,
Satisfiability Modulo Finite Fields \cite{SMFF}. As we
will show in subsequent sections, this interpretation supports
correctness guarantees, and also static type analyses for enforcing
confidentiality and integrity properties.

@@

We introduce the following syntax of SMT-style constraints
for $\minicat$:
$$
\begin{array}{rcl}
  \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs \mid \neg\eqs
\end{array}
$$
%
%In Figure \ref{fig-eqs} we define a syntax of SMT-style constraints
%$\eqs$, where
Note that constraint terms $\phi$ are similar to expressions $\be$
except that $\phi$ can ``cross party lines''. This is needed
to express correctness properties-- for example, in the
Goldreich-Micali-Wigderson (GMW) protocol wire values in circuits are
represented by reconstructive shares \cite{evans2018pragmatic}.  If by
convention shares of values $n$ are represented by $\mesg{n}$ on
clients, then assuming two clients $\setit{1,2}$ the reconstructed
value can be expressed as $\mx{n}{1} + \mx{n}{2}$.  So, while summing
values across clients is disallowed in $\minicat$ protocols, we can
express properties of shared values via constraints.

@@

\begin{fpfig}[t]{Interpretation of $\minicat$ expressions (T) and programs (B) as
  constraints}{fig-toeq}
\small{
\begin{mathpar}
  \toeq{x} = x

@@

  \toeq{\prog_1;\prog_2} = (\toeq{\prog_1} \wedge \toeq{\prog_2})
\end{mathpar}}
\end{fpfig}

@@

The mapping $\toeq{\cdot}$ from programs $\prog$ to constraints is
defined in Figure \ref{fig-toeq}. The interpretation of OT is omitted
from this figure which is general. For $\mathbb{F}_2$ (where we allow
negation $\neg$ of expressions) the interpretation is:
\begin{mathpar}
  \toeq{\elab{\OT{\be_1}{\cid_1}{\be_2}{\be_3}}{\cid_2}} =
  (\toeq{\elab{\be_1}{\cid_1}} \ftimes \toeq{\elab{\be_3}{\cid_2}}) \fplus
  (\neg\toeq{\elab{\be_1}{\cid_1}} \ftimes \toeq{\elab{\be_2}{\cid_2}}) 
\end{mathpar}
In general we will assume that any preprocessing predicate is defined
via a constraint $\eqspre$, i.e., $\preproc(\store) \iff \store
\models \eqspre$ for all $\store$.  The correctness of the
SMT interpretation of programs is characterized as follows.
\begin{theorem}
  \label{theorem-toeq}
  $\store$ is a model of $\eqspre \wedge \toeq{\prog}$ iff $\store \in \runs(\prog)$.
\end{theorem}

\subsubsection{Example: Correctness of 3-Party Addition}
Consider the following 3-party additive secret sharing protocol, where
each party generates two additive secret shares
\cite{evans2018pragmatic} for their respective secrets. Each party
publicly reveals the sums all their shares, and the sum of those is
the protocol output for each party.  {\footnotesize
$$
\begin{array}{c@{\qquad}c}
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
  \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
\end{array} & 
\begin{array}{lll}
  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\
  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
\end{array}
\end{array}
$$}
Letting $\prog$ be this protocol, we can verify correctness
as follows-- that is, the output sum of all shares is the sum of the
three input secrets:
$$
\toeq{\prog}\ \models\ \out{1} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
$$

>>>>

\section{$\minicat$ SMT-Based Hoare Logic}
\label{section-smt}

In previous work it was shown that the semantics of $\minicat$ in
$\mathbb{F}_2$ can be implemented with Datalog
\cite{skalka-near-ppdp24}, which is a form of constraint
programming. In this paper we extend this idea to arbitrary prime
fields by using a more general form of SMT constraint programming,
Satisfiability Modulo Finite Fields \cite{SMFF}. As we will show in
subsequent sections, this interpretation supports automated
enforcement of correctness guarantees, and also static type analyses
for enforcing confidentiality and integrity properties.
In this section we formulate a Hoare Logic for $\minicat$
as a foundation for protocol verification.

@@

We introduce the following syntax of SMT-style constraints for
$\minicat$:
$$
\begin{array}{rcl}
  \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs \mid \eqs \vee \eqs \mid \neg\eqs \mid \eqs \impl \eqs
\end{array}
$$
%
%In Figure \ref{fig-eqs} we define a syntax of SMT-style constraints
%$\eqs$, where
Note that constraint terms $\phi$ are similar to expressions $\be$
except that $\phi$ can ``cross party lines''. This is needed
to express correctness properties-- for example, in the
Goldreich-Micali-Wigderson (GMW) protocol wire values in circuits are
represented by reconstructive shares \cite{evans2018pragmatic}.  If by
convention shares of values $n$ are represented by $\mesg{n}$ on
clients, then assuming two clients $\setit{1,2}$ the reconstructed
value can be expressed as $\mx{n}{1} + \mx{n}{2}$.  So, while summing
values across clients is disallowed in $\minicat$ protocols, we can
express properties of shared values via constraints.

@@

\begin{fpfig}[t]{Interpretation of $\minicat$ expressions (T) and programs (B) as
  constraints}{fig-toeq}
\begin{mathpar}
  \toeq{x} = x

@@

  \toeq{\prog_1;\prog_2} = (\toeq{\prog_1} \wedge \toeq{\prog_2})
\end{mathpar}
\end{fpfig}

@@

The mapping $\toeq{\cdot}$ from programs $\prog$ to constraints is
defined in Figure \ref{fig-toeq}. The interpretation of OT is omitted
from this figure which is general. For $\mathbb{F}_2$ (where we allow
negation $\neg$ of expressions) the interpretation is:
\begin{mathpar}
  \toeq{\elab{\OT{\be_1}{\cid_1}{\be_2}{\be_3}}{\cid_2}} =
  (\toeq{\elab{\be_1}{\cid_1}} \ftimes \toeq{\elab{\be_3}{\cid_2}}) \fplus
  (\neg\toeq{\elab{\be_1}{\cid_1}} \ftimes \toeq{\elab{\be_2}{\cid_2}}) 
\end{mathpar}
In general we will assume that any preprocessing predicate is defined
via a constraint $\eqspre$, i.e., $\preproc(\store) \iff \store
\models \eqspre$ for all $\store$. Intuitively, a protocol satisfies
a correctness (post)condition $\eqs$ if its SMT interpretation
entails $\eqs$.

%The correctness of the
%SMT interpretation of programs is characterized as follows.
%\begin{theorem}
%  \label{theorem-toeq}
%  $\store$ is a model of $\eqspre \wedge \toeq{\prog}$ iff $\store \in \runs(\prog)$.
%\end{theorem}

\subsubsection{Example: Correctness of 3-Party Addition}
Consider the following 3-party additive secret sharing protocol, where
each party generates two additive secret shares
\cite{evans2018pragmatic} for their respective secrets. Each party
publicly reveals the sums all their shares, and the sum of those is
the protocol output for each party.  
$$
\begin{array}{c@{\qquad}c}
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
  \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
\end{array} & 
\begin{array}{lll}
  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\
  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
\end{array}
\end{array}
$$
Letting $\prog$ be this protocol, we can verify correctness
as follows-- that is, the output sum of all shares is the sum of the
three input secrets:
$$
\toeq{\prog}\ \models\ \out{1} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
$$

\noindent We can establish a logical identity between the models of a program and
its (non-adversarial) runs as follows.   
\begin{theorem}
  \label{theorem-toeq}
  $\sigma \models \eqspre \wedge \toeq{\prog}$ iff $\store \in \runs(\prog)$.
\end{theorem}
\noindent This result follows as a consequence of a more general Lemma
that incorporates a notion of preconditions. 
\begin{lemma}
  \label{lemma-toeq}
  Given $\eqs, \prog, \store$ with $\store \models \eqs$.
  Then $(\store,\prog) \redxs \store'$ iff $\store' \models
  \eqs \wedge \toeq{\prog}$.
\end{lemma}
\cnote{Prove me?}

\subsection{$\minicat$ Hoare Logic}

When reasoning about MPC protocols it is important to reason about
postconditions, e.g., to verify correctness conditions. It is also
important to reason about preconditions of protocols, e.g., to express
properties of assumed initialization protocols such as Beaver Triples.
To support this reasoning we introduce a Hoare logic for $\minicat$,
where $\htrip{\eqs_1}{\prog}{\eqs_2}$ means that $\eqs_1,\eqs_2$ are
the preconditions and postconditions, respectively, of $\prog$.
A semantic notion of validity for Hoare triples is defined as
follows:

\minicathtripfig

\begin{definition}
$\htrip{\eqs_1}{\prog}{\eqs_2}$ is \emph{valid} iff $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$.
\end{definition}

The following Lemma establishes that this is a meaningful (and
standard) definition. It follows by Lemma \ref{lemma-toeq}
and transitivity of $\models$.
\begin{lemma}
  \label{lemma-hoare}
  Given $\eqs_1, \eqs_2, \prog, \store$ with $\store \models \eqs_1$
  and $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$. Then
  $(\store,\prog) \redxs \store'$ implies $\store' \models
  \eqs_2$.
\end{lemma}

Now we can introduce a validity-preserving derivation system for the
$\minicat$ Hoare logic. Since the language is computationally simple
the Hoare logic can be relatively simple and mostly standard.
The derivation rules are specified in Figure \ref{fig-minicathtrip}.
They are mostly familiar except perhaps for \TirName{Assert}, which
just enforces the expected equality as a precondition, and
\TirName{Hyp}. It is unknown to the authors and not immediately
obvious whether \TirName{Hyp} is derivable from other rules, but
it is validity preserving as we show below. Its main use in this
paper will be to enable an automated Hoare logic check we develop
later (Section \ref{section-metahtrip}).
\begin{lemma}
  If $\htrip{\eqs_1}{\prog}{\eqs_2}$ is derivable, then it is valid.
\end{lemma}

\begin{proof}
  Assume $\htrip{\eqs_1}{\prog}{\eqs_2}$ is derivable without \TirName{Hyp}.
  Then $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$, since all the other rules
  are well-known to be validity preserving. 
  Now since $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$ by assumption, we have
  $
  \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 
  $,
  and clearly:
  $$
  \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 \wedge (\eqs_2 \impl \eqs)
  $$
  and since $\eqs_2 \wedge (\eqs_2 \impl \eqs)$ iff
  $\eqs_2 \wedge \eqs$, we have
  $\eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 \wedge \eqs$.
\end{proof}
